---
title: "Create Survival Data"
output: pdf_document
---

```{r setup, include=FALSE}
# Set the default CRAN repository
local({r <- getOption("repos")
       r["CRAN"] <- "http://cran.r-project.org" 
       options(repos=r)
})

# Set code chunk defaults
knitr::opts_chunk$set(echo = TRUE)

# Load required packages; install if necessary
# CAUTION: DO NOT interrupt R as it installs packages!!
if (!require("ggplot2")) {
  install.packages("ggplot2")
  library(ggplot2)
}

if (!require("knitr")) {
  install.packages("knitr")
  library(knitr)
}
if (!require("readr")) {
  install.packages("readr")
  library(readr)
}
if (!require("dplyr")) {
  install.packages("dplyr")
  library(dp)
}

if (!require("RColorBrewer")) {
  install.packages("RColorBrewer")
  library(RColorBrewer)
}
if (!require("beeswarm")) {
  install.packages("beeswarm")
  library(beeswarm)
}
if (!require("tidyverse")) {
  install.packages("tidyverse")
  library(tidyverse)
}
if (!require("ggbeeswarm")) {
  install.packages("ggbeeswarm")
  library(ggbeeswarm)
}
if (!require("xts")) {
  install.packages("xts")
  library(xts)
}
if (!require("plotly")) {
  install.packages("plotly")
  library(plotly)
}
if(!require("lubridate")) {
  install.packages("lubridate")
  library(lubridate)
}
if(!require("survival")) {
  install.packages("survival")
  library(survival)
}
if(!require("survminer")) {
  install.packages('survminer')
  library(survminer)
}
if(!require("ranger")){
  install.packages("ranger")
  library(ranger)
}
if(!require("ggfortify")){
  install.packages("ggfortify")
  library(ggfortify)
}

```

This notebook shows how to use the transaction data from a deployment of AAVE and turn it into a set of survival data for use in survival analysis.

## Load the data from the IDEA server:

First, we define the parts of the filepaths for loading the transaction data from the IDEA server:

```{r}
dataPath = "/data/IDEA_DeFi_Research/Data/Lending_Protocols/Aave/"
V2Path = paste0(dataPath, "V2/")
V3Path = paste0(dataPath, "V3/")
transactionsFile = "transactions.csv"
clustersFile = "userClusters.csv"
marketTypesFile = "marketTypes.csv"
```

```{r}
allTransactions <- read_csv(paste0(V2Path, "Mainnet/", transactionsFile))
```

Next we do a bit of pre-processing, setting up a couple of helper functions and dataframes to give us some additional information about the transactions that will be useful when using the survival analysis data.

```{r}
# For consistency's sake, because we currently have to manually update the datasets and therefore there's no guarantee that they run to the same date,
#   let's choose a cutoff date that remove data past that point:
cutoffDate = "2022-10-01 00:00:00 UTC"

coinStabilities <- allTransactions %>%
  select(reserve) %>%
  distinct() %>%
  mutate(coinType = case_when(reserve %in% c("USDT", "USDC", "TUSD", "DAI", "SUSD", "USDT.e", "DAI.e", "USDC.e", "BUSD", "GUSD", "LUSD", "USDt", "1DAI", "1USDT", "1USDC", "fUSDT") ~ "Stable",
                              TRUE ~ "Non-Stable"))

not_all_na <- function(x) any(!is.na(x))
`%notin%` <- Negate(`%in%`)

fullDF <- allTransactions %>%
  mutate(datetime = as_datetime(timestamp)) %>%
  filter(datetime <= cutoffDate) %>%
  left_join(coinStabilities, by = "reserve") %>%
  mutate(quarter = floor_date(datetime, unit = "quarter")) %>%
  group_by(type) %>%
  mutate(amountUSDQuartile = ntile(amountUSD, n = 4),
         principalUSDQuartile = ntile(principalAmountUSD, n = 4),
         collateralUSDQuartile = ntile(collateralAmountUSD, n = 4)) %>%
  group_by(reserve) %>%
  mutate(amountQuartile = ntile(amount, n = 4)) %>%
  ungroup()

quarters <- fullDF %>%
    select(quarter) %>%
    distinct() %>%
    arrange(quarter)

users <- fullDF %>%
  select(user) %>%
  distinct()

survivalData <- NULL
```
```{r}
sampleProportion = 0.2

sampledUsers <- users %>%
  slice_sample(prop = sampleProportion)


df <- fullDF %>%
  filter(user %in% sampledUsers$user)

```


```{r}

for(timeStart in quarters$quarter){
  timeFinal = ceiling_date(as_datetime(timeStart+1), unit = "quarter")

  indexEvents <- df %>%
    filter(type == "borrow",
           quarter == timeStart) %>%
    select(where(not_all_na)) %>%
    mutate(ID = row_number(),
           indexTime = timestamp,
           indexAmountUSDQuartile = amountUSDQuartile,
           indexMarketType = marketType)
  
  outcomeEvents <- df %>%
    filter(type == "repay",
           quarter == timeStart) %>%
    select(where(not_all_na)) %>%
    mutate(outcomeTime = timestamp,
           outcomeAmountUSDQuartile = amountUSDQuartile,
           outcomeMarketType = marketType)
  
  indexToOutcomes <- left_join(indexEvents, outcomeEvents, by = c("user", "reserve")) %>% 
    group_by(ID) %>% 
    mutate(status = case_when(indexTime >= max(outcomeTime) ~ 0,
                            is.na(outcomeTime) ~ 0,
                            TRUE ~ 1)) %>%
    distinct()
  
  rightCensoredEvents <- indexToOutcomes %>%
    filter(status == 0) %>%
    group_by(ID) %>%
    summarise(user, timeDiff = as.double(timeFinal - indexTime), status, reserve, reserveType = coinType.x, timeStart, amountUSDQuartile = indexAmountUSDQuartile, marketType = indexMarketType) %>%
    distinct()
  
  uncensoredEvents <- indexToOutcomes %>%
    filter(status == 1) %>%
    filter(outcomeTime > indexTime) %>%
    summarise(user, timeDiff = min(outcomeTime) - indexTime, status, reserve, reserveType = coinType.x, timeStart, amountUSDQuartile = indexAmountUSDQuartile, marketType = indexMarketType) %>%
    distinct()
  
  leftCensoredEvents <- anti_join(outcomeEvents, indexEvents, by = c("user", "reserve")) %>%
    mutate(status = 0) %>%
    summarise(user, timeDiff = as.double(timestamp - timeStart), status, reserve, reserveType = coinType, timeStart, amountUSDQuartile = outcomeAmountUSDQuartile, marketType = outcomeMarketType)
  
  survivalData <- bind_rows(survivalData, rightCensoredEvents, uncensoredEvents, leftCensoredEvents)
}
```

```{r, fig.height = 10, fig.width=7.5}
survData <- survivalData %>%
  mutate(datetime = as_datetime(timeStart)) %>%
  filter(timeStart != max(quarters$quarter),
         timeStart != min(quarters$quarter)) %>%
  mutate(quarter = paste0(year(datetime), " Q", quarter(datetime)))

legendLabels <- survData %>%
  group_by(quarter) %>%
  summarise() %>%
  arrange(quarter)

# These colors were hand-picked from https://r-charts.com/colors/ with the attempt to align the colors with the seasons:
quarterColors <- c("#98F5FF", # cadetBlue1
                   "#FFBBFF", # plum1
                   "#00CD66", # springgreen3
                   "#CD4F39", # tomato2
                   "#7AC5CD", # cadetBlue3
                   "#CD96CD", # plum3
                   "#008B45"  # springgreen4
                   )

survFit <- survfit(Surv(timeDiff/86400, as.numeric(status)) ~ quarter, data=survData)

survPlot <- ggsurvplot(survFit, 
                      censor = FALSE,
                      title = "How Long Do Users Wait Before Making a Repayment?",
                      xlab = "Time (days)",
                      ylab = "Probability of No Repay",
                      legend = "right",
                      legend.labs = legendLabels$quarter,
                      legend.title = "Year: Quarter",
                      pval = TRUE,
                      palette = quarterColors,
                      risk.table = TRUE
                      )
survPlot
```

```{r, fig.height=10, fig.width = 7.5}
fit <- coxph(Surv(timeDiff / 86400, as.numeric(status)) ~ reserveType + quarter + amountUSDQuartile + marketType, data = survData)
ggcoxzph(cox.zph(fit))
```

```{r}
ggforest(fit, data = survData)
```

```{r}
survivalData <- NULL
for(timeStart in quarters$quarter){
  timeFinal = ceiling_date(as_datetime(timeStart+1), unit = "quarter")

  indexEvents <- df %>%
    filter(type == "deposit",
           quarter == timeStart) %>%
    select(where(not_all_na)) %>%
    mutate(ID = row_number(),
           indexTime = timestamp,
           indexAmountUSDQuartile = amountUSDQuartile,
           indexMarketType = marketType)
  
  outcomeEvents <- df %>%
    filter(type == "redeem",
           quarter == timeStart) %>%
    select(where(not_all_na)) %>%
    mutate(outcomeTime = timestamp,
           outcomeAmountUSDQuartile = amountUSDQuartile,
           outcomeMarketType = marketType)
  
  indexToOutcomes <- left_join(indexEvents, outcomeEvents, by = c("user", "reserve")) %>% 
    group_by(ID) %>% 
    mutate(status = case_when(indexTime >= max(outcomeTime) ~ 0,
                            is.na(outcomeTime) ~ 0,
                            TRUE ~ 1)) %>%
    distinct()
  
  rightCensoredEvents <- indexToOutcomes %>%
    filter(status == 0) %>%
    group_by(ID) %>%
    summarise(user, timeDiff = as.double(timeFinal - indexTime), status, reserve, reserveType = coinType.x, timeStart, amountUSDQuartile = indexAmountUSDQuartile, marketType = indexMarketType) %>%
    distinct()
  
  uncensoredEvents <- indexToOutcomes %>%
    filter(status == 1) %>%
    filter(outcomeTime > indexTime) %>%
    summarise(user, timeDiff = min(outcomeTime) - indexTime, status, reserve, reserveType = coinType.x, timeStart, amountUSDQuartile = indexAmountUSDQuartile, marketType = indexMarketType) %>%
    distinct()
  
  leftCensoredEvents <- anti_join(outcomeEvents, indexEvents, by = c("user", "reserve")) %>%
    mutate(status = 0) %>%
    summarise(user, timeDiff = as.double(timestamp - timeStart), status, reserve, reserveType = coinType, timeStart, amountUSDQuartile = outcomeAmountUSDQuartile, marketType = outcomeMarketType)
  
  survivalData <- bind_rows(survivalData, rightCensoredEvents, uncensoredEvents, leftCensoredEvents)
}
```

```{r, fig.height=10, fig.width=7.5}
survData <- survivalData %>%
  mutate(datetime = as_datetime(timeStart)) %>%
  filter(timeStart != max(quarters$quarter),
         timeStart != min(quarters$quarter)) %>%
  mutate(quarter = paste0(year(datetime), " Q", quarter(datetime))) %>%
  select(-datetime, -timeStart)

legendLabels <- survData %>%
  group_by(quarter) %>%
  summarise() %>%
  arrange(quarter)

# These colors were hand-picked from https://r-charts.com/colors/ with the attempt to align the colors with the seasons:
quarterColors <- c("#98F5FF", # cadetBlue1
                   "#FFBBFF", # plum1
                   "#00CD66", # springgreen3
                   "#CD4F39", # tomato2
                   "#7AC5CD", # cadetBlue3
                   "#CD96CD", # plum3
                   "#008B45"  # springgreen4
                   )

survFit <- survfit(Surv(timeDiff/86400, as.numeric(status)) ~ quarter, data=survData)

survPlot <- ggsurvplot(survFit, 
                      censor = FALSE,
                      title = "How Long Do Users Leave Funds In Their Account?",
                      xlab = "Time (days)",
                      ylab = "Probability of No Withdrawal",
                      legend = "right",
                      legend.labs = legendLabels$quarter,
                      legend.title = "Year: Quarter",
                      pval = TRUE,
                      palette = quarterColors,
                      risk.table = TRUE
                      )
survPlot
```

```{r, fig.height=10, fig.width = 7.5}
fit <- coxph(Surv(timeDiff / 86400, as.numeric(status)) ~ quarter, data = survData)
ggcoxzph(cox.zph(fit))
```

```{r, fig.width=7.5, fig.height=7.5}
ggforest(fit, data = survData, fontsize = 1.1, main = "How Does Time to Repayment Compare Between Quarters?")
```

```{r}
averageLoanSizesByQuarter <- fullDF %>%
  filter(type == "borrow") %>%
  dplyr::group_by(quarter) %>%
  dplyr::summarize(meanLoanSize = mean(amountUSD))

averageDepositSizesByQuarter <- fullDF %>%
  filter(type == "deposit") %>%
  dplyr::group_by(quarter) %>%
  dplyr::summarize(meanDepositSize = mean(amountUSD))
```